---
title: "Getting started with reread"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with reread}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

`reread` detects careless or insufficient-effort responding in questionnaire
data. This vignette walks through the one-call workflow, the pieces underneath,
and how to compare against other indices.

```{r setup}
library(reread)
```

## One call

The main function takes a matrix or data frame of Likert responses (respondents
in rows, items in columns) and returns scores, flags, and a calibrated
prevalence estimate.

```{r detect}
fit <- reread(demo_careless$responses)
fit
```

The per-respondent output is a data frame:

```{r scores}
head(fit$scores)
```

`rc` is the coherence index (low = careless); `eta` is the ensemble log-odds
(high = careless); `prob` is the careless probability; `flagged` is the decision.
Because `demo_careless` is simulated, we can check the flags against the truth:

```{r truth}
table(flagged = fit$flagged, careless = demo_careless$careless)
```

## Sensitivity

There is one shipped cut and normally nothing to choose: it sits 2.5 standard
deviations above the attentive mode and is the better setting from the cleanest
samples up to a careless rate of about 35%. Past that a more liberal cut recovers
more of the careless, and you ask for it explicitly:

```{r custom-z}
fl <- auto_flag(ensemble_score(demo_careless$responses)$eta, z = 1.5)
fl$n_flagged
```

If you already know (or can estimate) the careless rate, supply it and the
automatic threshold is bypassed:

```{r manual}
reread(demo_careless$responses, prevalence = 0.2)$n_flagged
```

## Under the hood

The ensemble score is a reliability-gated logistic combination of `rc`,
longstring, and person--total, with frozen weights:

```{r pieces}
es <- ensemble_score(demo_careless$responses)
str(es[c("gate", "profile_info", "signal_strength")])
```

The label-free threshold comes from a one-sided fit of the attentive mode on the
ensemble log-odds: the dominant mode `m`, a scale read from that mode's left
flank alone, and a cut at `m + 2.5 * sigma`. The rule is applied
unconditionally, so a careless-free sample is not guaranteed to yield nothing —
it yields its own 2.5-sigma tail, on the order of one to two per cent of
respondents. That cost is a property of the sample's size, not a surprise:
report it alongside the flags rather than assuming it away.

```{r flag}
fl <- auto_flag(es$eta)
c(flagged = fl$n_flagged, prevalence = round(fl$prevalence, 3), status = fl$status)
```

Because the rule is equivariant to the scale of whatever score it is given, it
can be applied to a single index instead of the ensemble. On external data the
ensemble and the person-total correlation are close to interchangeable, so that
choice is offered rather than hidden — though the ensemble remains the default,
being the more robust of the two where the ground truth is clean:

```{r pt}
pt <- reread(demo_careless$responses, score = "person_total")
c(ensemble = sum(fit$flagged), person_total = pt$n_flagged)
```

## Benchmarking

`benchmark_indices()` compares the ensemble against the auxiliary indices by AUC,
with DeLong confidence intervals and tests (requires `pROC`):

```{r bench, eval = requireNamespace("pROC", quietly = TRUE)}
benchmark_indices(demo_careless$responses, demo_careless$careless)
```

## Simulating your own data

```{r sim}
clean <- simulate_clean(n_factors = 10, items_per_factor = 6, n = 200, seed = 1)
dat   <- inject_careless(clean$data, prevalence = 0.2, seed = 2)
reread(dat$data)$n_flagged
```

## Scope

`reread` targets *inconsistent* careless responding on multi-construct
questionnaires (roughly 60+ items). It warns on data with no usable structure,
does not detect content-responsive distortions such as faking, and on short or
single-factor scales leans on the auxiliary detectors.
